Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-04-09 manuscript.qmd
change to manuscript type & move config

---
title: "Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer"
authors:
  - name: Roman Ludwig
    orcid: 0000-0001-9434-328X
    email: roman.ludwig@usz.ch
  - name: Yoel Perez Haas
    email: yoel.perezhaas@usz.ch
  - name: Jan Unkelbach
    orcid: 0000-0002-4275-990X
    email: jan.unkelbach@usz.ch
bibliography: references.bib
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
  pdf:
    keep-tex: true
    include-in-header:
    - text: |
        \usepackage{multirow}
        \usepackage{centernot}
jupyter: python3
execute: 
  cache: true
crossref:
  fig-prefix: "figure"
  tbl-prefix: "table"
  eq-prefix: "equation"
  sec-prefix: "section"
code-links:
  - text: GitHub Repo
    href: https://github.com/rmnldwg/bilateral-paper
    icon: github
resources: 
  - data/reduced.csv
meca-bundle: true
---

@@

- blabla
- head and neck cancer spreads through the lymphatic network
- may sometimes spread to contralateral side
- spreads more frequently and to larger extent contralaterally when tumor extends the mid-sagittal line
- we describe a model based on previously published hidden Markov model
- we extend it to cover the contralateral side, too
- naive approach: make two independent models for each side, but that is not supported by the data
- ipsi- and contralateral side are correlated via time of diagnosis, which is correlated with T-category
- tumor extension over mid-sagittal line is modelled as random variable
- spread probabilities from tumor to contralateral LNLs in case of midline extesnion are linear combinations of these probabilities in case of ipsilateral spread and contralateral spread when tumor is clearly lateralized

>>>>

---
title: "Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer"
---

@@

- head and neck cancer spreads through the lymphatic network
- may sometimes spread to contralateral side
- spreads more frequently and to larger extent contralaterally when tumor extends the mid-sagittal line
- we describe a model based on previously published hidden Markov model
- we extend it to cover the contralateral side, too
- naive approach: make two independent models for each side, but that is not supported by the data
- ipsi- and contralateral side are correlated via time of diagnosis, which is correlated with T-category
- tumor extension over mid-sagittal line is modelled as random variable
- spread probabilities from tumor to contralateral LNLs in case of midline extesnion are linear combinations of these probabilities in case of ipsilateral spread and contralateral spread when tumor is clearly lateralized
